Modelling the Retinotectal projection - can competition provide a stopping mechanism?

### 2022-02-08 paper/paper.tex
WIP

% G - chemoaffinity description
%\subsection{Chemoaffinity}
The chemoaffinity component, $\mathbf{G}_b$, is governed by receptor-ligand
signalling. The signal transmitted when a tectal ligand binds to a RGC
receptor on a branch (forward signalling) can lead to one of two effects; the
signal may induce an attraction towards the ligand-expressing region or a
repulsion away from it.
%
For axon-tectum interactions, we implemented attraction as a climbing of the
tectal ligand expression gradient and repulsion as gradient descent.
%
To determine the strength of the effect we assumed a purely linear receptor
binding model, and set the chemotactic movement vector of the branch $b$ at
location $\mathbf{x}_b$ on the tectum to be
%
\begin{equation}
\mathbf{G}_b = \sum_i^N F_i\,r_{i,b} \nabla L_i(\mathbf{x}_b)
\end{equation}
%
where $r_{i,b}$ is the receptor expression on branch $b$ for ligand-receptor
pair $i$, $\nabla L_i$ is the gradient of expression of ligand $i$ on the
tectum and $F_i$ denotes the direction of the interaction induced when a
molecule of ligand $i$ binds to a receptor $i$ molecule. $F_i$ takes the value
$-1$ for a repulsive interaction or $1$ for an attractive interaction.
%
We assumed that all
receptor-ligand signalled interactions are repulsive ($F_i=-1, \forall i$), as
for EphA-ephrinA
coupling \citep{drescher_vitro_1995,nakamoto_topographically_1996}.
%
$m_{\!_G}$ in Eq.\,\ref{e:mv} is a
scalar parameter which controls how much movement is generated for a given
chemotactic movement vector, $\mathbf{G}_b$.
% end chemoaffinity description

@@

>>>>

% G - chemoaffinity description
%\subsection{Chemoaffinity}
The chemoaffinity component, $\mathbf{G}_b$, is governed by receptor-ligand
signalling. The signal transmitted when a tectal ligand binds to a RGC
receptor on a branch (forward signalling) can lead to one of two effects; the
signal may induce an attraction towards the ligand-expressing region or a
repulsion away from it.
%
For axon-tectum interactions, we implemented attraction as a climbing of the
tectal ligand expression gradient and repulsion as gradient descent.
%
To determine the strength of the effect we assumed a purely linear receptor
binding model, and set the chemotactic movement vector of the branch $b$ at
location $\mathbf{x}_b$ on the tectum to be
%
\begin{equation}\label{e:G}
\mathbf{G}_b = \sum_i^N F_i\,r_{i,b} \nabla L_i(\mathbf{x}_b)
\end{equation}
%
where $r_{i,b}$ is the receptor expression on branch $b$ for ligand-receptor
pair $i$, $\nabla L_i$ is the gradient of expression of ligand $i$ on the
tectum and $F_i$ denotes the direction of the interaction induced when a
molecule of ligand $i$ binds to a receptor $i$ molecule. $F_i$ takes the value
$-1$ for a repulsive interaction or $1$ for an attractive interaction.
%
We assumed that all
receptor-ligand signalled interactions are repulsive ($F_i=-1, \forall i$), as
for EphA-ephrinA
coupling \citep{drescher_vitro_1995,nakamoto_topographically_1996}.
%
$m_{\!_G}$ in Eq.\,\ref{e:mv} is a
scalar parameter which controls how much movement is generated for a given
chemotactic movement vector, $\mathbf{G}_b$.
% end chemoaffinity description

@@

\subsection*{Robustness to noise}

\emph{This section might fit most naturally after the Chemoaffinity only section.}

\begin{figure}
\includegraphics[width=0.3\linewidth]{./images/GI_vs_noise_tweaked.png}
\caption{The mean value of $\eta$ (number of crossings) and $\epsilon$
(mean distance error), averaged over the period from simulation time $t=500$ to $t=1000$, plotted with respect to the
noise gain in the $\mathbf{GI}$ model and the wildtype condition (compare with
Fig.\,\ref{f:GI}). Both metrics indicate that a moderate
amount of gradient sampling noise is \emph{beneficial} to the establishment of
an ordered tectum.}
\label{f:noise}
\end{figure}

One problem with axon gradient-following is the ability of the axon to
successfully detect the direction of the gradient given the small distances
over which the axon can sample signalling molecules and the size of receptor
binding noise \citep{goodhill_noise_2014}.  We investigated the ability of the
system to self-organise in the presence of noise, by introducing a stochastic
element to each branch's sampling of the gradient.

% See branch::estimate_ligand_gradient() in sim/agent/branch.h for the code
Noise is added to the estimation of the ligand gradient by multiplying the
true gradient by a random vector, $\mathbf{U}$, whose coordinates are drawn
from a random uniform distribution centred about 0 with magnitude $\nu$, denoted:
$\mathbf{U}(-\nu, \nu)$. Eq.\,\ref{e:G} becomes

\begin{equation}\label{e:Gnu}
\mathbf{G}_b = \sum_i^N F_i\,r_{i,b}\, \mathbf{U}(-\nu, \nu) \nabla L_i(\mathbf{x}_b)
\end{equation}

Fig.\,\ref{f:noise} shows the response of the $\mathbf{GI}$ model to the
addition of ligand gradient estimation noise.

What about retinal expression noise? This would have a disordering effect on
the pattern.
